% ************************** Thesis Acknowledgements *****************************

\begin{acknowledgements}

\noindent Thank you to my supervisors Nick Wright, Stephen McGough, and Per Berggren for continued support and expert advice throughout my PhD. \newline

\noindent Thank you to Paul Watson and Darren Wilkinson for giving me the opportunity to pursue a PhD through the Cloud Computing for Big Data CDT. \newline

\noindent Thank you to Matthew Sharpe, Georgia Atkinson, and Kirsten Crane for your aid during fieldwork and understanding the frustrations of working with messy ecology data.\newline

\noindent Thank you to Barbara Cheney (University of Aberdeen), M\`{o}nica Arso Civil (University of St Andrews), and Reny Tyson Moore and Jason Allen (Chicago Zoological Society's Sarasota Dolphin Research Program) for providing additional photo-id data for use in this thesis.\newline

\noindent Thank you to the citizen scientists of both the Newbiggin-by-the-Sea Dolphin Watch and the North East Cetacean Project. Your dedication was vital to the success of my fieldwork. \newline

\noindent Thank you to Jennifer Wood and Andrew Turnbull for your indispensable aid in navigating the endless minefield that is University administration.\newline

\noindent Thank you to those who have provided engaging discussions (and distractions) throughout my PhD, including but not limited to: Tom Owen, Tom Lowe, Kieran Peel, Jack Aiston, Dale Whinham, Phillip Darke, and Carlos Vladimiro González Zelaya. \newline

\noindent Finally, thank you to my parents, Paula and Norman Trotter. I would not have achieved what I have without you.\newline

\noindent \centering\textit{This work was supported by the Engineering and Physical Sciences Research Council Centre for Doctoral Training in Cloud Computing for Big Data\newline[EP/L015358/1]} 

\end{acknowledgements}
